---
title: "Evaluation_techniques"
output: html_document
date: "2023-10-01"
---
#Installing packages
```{r}
#install.packages("dplyr")
#install.packages("tidyverse")
#install.packages("yarrr")

library(dplyr)
library(tidyverse)
library(yarrr)

#One goal that I have this semester is to be more confident in my R skills. 
```

#Getting Started with R
```{r}
getwd()
setwd(dir = "/Users/oliviaenriquez/Desktop/Fall_23/Program_evaluation/")
```

```{r}
data <- read.csv("faketucky.csv", na = "999")
#data <- read_excel("my_file.xlsx")
```

```{r}
head(pirates)
#view(pirates)
summary(pirates$tattoos)
summarize

```

```{r}
min(data$gpa)
min(pirates$parrots)

max(data$gpa)
mean(pirates$parrots)

unique(pirates$college)
```

```{r}
aggregate(x = parrots ~ sex,
          data = pirates,
          FUN = mean)
```

How many school districts are in the data?
Hint: use unique() and length() function

What is the median GPA?
Hint: make sure missing values are NA

What percent of students had perfect attendance?
Hint: use nrow(data[data$column == 0, ])

Upload a screenshot of your code and output.

```{r}
length(unique(data$school_district))
```

```{r}
summary(data$gpa)
median(data$gpa, na.rm = TRUE)
```

```{r}
summary(data$percent_absent)

perfect_attendance <- nrow(data[data$percent_absent ==0, ])
num_students<- length(data$percent_absent)

(perfect_attendance-111)/num_students*100

data$perfect_attendance[data$percent_absent == 0] <-1
summary(data$perfect_attendance)
table(data$perfect_attendance)

```


Answer the following questions using the faketucky.csv dataset.
1. Change one variable to a factor.
2. Conduct a t-test between the gpa of boys and girls. What is the mean of each group?
3. Conduct a regression (linear model) with gpa as the x variable and male as the y variable. Is the result significant? How do you know? What are the limitations of this model?

```{r}
#is.factor(faketucky$male)
#faketucky$male <- as.factor(faketucky$male)
#is.factor(faketucky$male)
#class(faketucky$male)
```
```{r}
faketucky <- read.csv("faketucky.csv", na = "999") 
summary(faketucky$gpa)
```

```{r}
options(scipen=999)
t.test(faketucky$gpa ~ faketucky$free_and_reduced_lunch)
```

#Bivariate and Multivariate Regressions
```{r}
model <- lm(faketucky$gpa ~ faketucky$free_and_reduced_lunch)

model2 <- lm(faketucky$gpa ~ faketucky$free_and_reduced_lunch + faketucky$male + faketucky$race_ethnicity)

summary(model)
summary(model2)
```

##Stargazer model
```{r}
#install.packages("stargazer")
library(stargazer)

#as.factor(faketucky$race_ethnicity)
stargazer(model, type="text", title="Regression Results", header=FALSE, keep.stat="n", single.row=TRUE, digits =2, out = "regression.results.doc")

```


#Difference in Difference, ggplot
```{r}
data <- data %>%
  mutate(year = sample(2020:2021, nrow(data), replace = TRUE))
```

```{r}
data$year <- factor(data$year, levels = c("2020", "2021"))
levels(data$year)
```


```{r}
data$treated <- ifelse(data$first_high_school_attended == "Jackson", 1, 0)
```

```{r}
tapply(data$gpa, list(data$treated, data$year), mean, na.rm = TRUE)
```

```{r}
school_gpa_model <- lm(data$gpa ~ data$treated * data$year)
summary(school_gpa_model)
```

```{r}
library(ggplot2)

# Calculate the means by group and time variable
plot_data <- aggregate(gpa ~ treated + year, data, mean)
```

```{r}
# Create a DiD plot
ggplot(plot_data, aes(x = as.numeric(year), y = gpa, color = factor(treated))) +
  geom_line() +                           # Line plot for each group
  geom_smooth(method = "lm", se = FALSE) + # Add regression lines
  labs(title = "Difference-in-Differences Plot",
       x = "Year",
       y = "GPA") +
  scale_color_manual(values = c("lightblue", "pink"), 
                     labels = c("Non-Program Participants", "Program Participants")) +
  scale_x_discrete(limits = c("Pre-program", "Post-program")) +
  theme_minimal()

```

#Random Assignment
```{r}
library(dplyr)

data <- data %>%
  filter(grepl("Jackson|Wallaby", first_high_school_attended))

table(data$first_high_school_attended)

data$participants <- ifelse(data$first_high_school_attended == "Jackson", 1, 0)
data$participants <- as.factor(data$participants)
```
#Propensity Score Matching
```{r}
library(MatchIt)

data <- data[!is.na(data$gpa), ]
data <- data[!is.na(data$free_and_reduced_lunch), ]

m.nn <- matchit(participants ~ male + race_ethnicity + free_and_reduced_lunch + percent_absent + gpa,  data = data, method = "nearest", ratio = 1)

summary(m.nn)

plot(m.nn, type = "jitter")
plot(m.nn, type = "hist")
```


```{r}
data_matched<-match.data(m.nn)
```

```{r}
model <- lm(received_high_school_diploma ~ participants, data = data_matched)
summary(model)
```

```{r}
model2 <- lm(gpa ~ participants, data = data_matched)
summary(model2)
```










